---
title: "412 Project"
author: "Tianshu Fan, Jaehee Jeong, Lisa Kaunitz"
date: "11/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(DataExplorer)
library(lubridate)
library(caTools)
library(MASS)
library(leaps)
library(caret)
library(pcr)
library(pls)
library(Metrics)
```

# Import Data

```{r}
house <- read_csv("/Users/jaeheejeong/Desktop/UCLA/412-Advanced Regression and Predictive Modeling/Project/STAT_412/kc_house_data.csv")
# View(house)

```


Variables with the highest correlation with `price`: 
- sqft_living
- grade
- sqft_above
- sqft_living15
- bathrooms

Added 5 more variables
- lat, bedrooms, floors, sqft_basement, view

# Cleaning Data

```{r}
# remove outliers for Price 
IQR <- 645000-321950
Upper <- 1.5*IQR + 540088
Lower <- 540088 - 1.5*IQR
house <- subset(house, price >= Lower & price <= Upper)
summary(house)

# Add a feature if there is a basement then 1 else 0
for(i in 1: nrow(house)){
    if (house$sqft_basement[i] >0) {
  house$sqft_basement_yesno[i] <- 1
  } else {
  house$sqft_basement_yesno[i] <- 0
  }
}

# drop id
house<-house[,-1]

# tbd: as.factor (depending on random forest output)
```

# EDA 

```{r}
#create_report(house)
summary(house)
# calculate standard deviation
round(sapply(house, sd),2)

ggplot(house, aes(x=date, y = price))+
  geom_line()+
  xlab('Date')+
  ylab('Price')+
  ggtitle('House Prices Over Time') + 
  theme_bw()

# Distribution of price
ggplot(house)+
  aes(x=price)+
  geom_histogram(col = 'black', bins = 20) +
  ggtitle("Distribution of Price (no transformation)") +
  theme_bw()

# log transform 
house$log_price <- log(house$price)
ggplot(house)+
  aes(x=log_price)+
  geom_histogram(col = 'black', bins = 20) +
  ggtitle("Distribution of Price (log transformation)") +
  theme_bw()

#grade, sqft_living, sqft_living15, sqft_above, bathrooms, lat, bedrooms, floors, sqft_basement, view. 

# Untransformed price
pairs(house[,c(11,5,19,12,4,17,3,7,13,9)], pch = 19, lower.panel = NULL)

# Log Price
pairs(house[,c(11,5,19,12,4,17,3,7,13,9)], pch = 19, lower.panel = NULL)


ggplot(house[1:100,], aes(long, lat), color = "grey99") +
  borders("state") + geom_point() +
  ggrepel::geom_label_repel(aes(label = price)) +
  theme_void() + coord_cartesian()

usa <- map_data("usa")

wa_df <- subset(states, region == "washington")
washington_base <- ggplot(data = ca_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

states <- map_data("state")
```


# Feature Importance (Random Forest)

```{r}
# TianShu 


```


# Data Cleaning (post feature importance)

```{r}
# Drop ID, Date, Zipcode
drop <- c('date','id','zipcode')
house <- house[!names(house) %in% drop]

# Factor variables (tbd)
house$bedrooms = as.factor(house$bedrooms)
house$bathrooms= as.factor(house$bathrooms)
house$floors = as.factor(house$floors)
house$waterfront = as.factor(house$waterfront)
house$view = as.factor(house$view)
house$condition = as.factor(house$condition)
house$grade = as.factor(house$grade)

```

# Splitting Data (25-75)

```{r}
split <- createDataPartition(house$price, p=0.75, list = F)
train <- house[split,]
test <- house[-split,]
```

# Linear Regression Model

Start off with the most basic model

```{r}

LinearReg_1 <- lm(log_price ~ bedrooms + bathrooms + sqft_living + sqft_lot + floors + 
  waterfront + view +  + grade + yr_built + yr_renovated + 
  zipcode + lat + long + sqft_living15 + sqft_above, data = train)

summary(LinearReg_1)
# yr_renovated, sqft_above are not stat sig.
# R_squared = 0.77
plot(LinearReg_1)
# meets model assumptions. 

LinearReg_1_RSS <- c(crossprod(LinearReg_1$residuals)) # RSS = 1026.547
LinearReg_1_MSE <- LinearReg_1_RSS / length(LinearReg_1$residuals) # MSE = 0.06332801
LinearReg_1_RMSE <- sqrt(LinearReg_1_MSE) # RMSE = 0.2516506

```

__Linear Regression Results__: 

R_squared = 0.77
RSS = 1026.547
MSE = 0.06332801
RMSE = 0.2516506



## Forward & Backward Selection 

```{r}
Step_1 <- stepAIC(LinearReg_1, direction = "both")
Step_1$anova

# Telling us to get rid of `sqft_above`
```

# PCA

```{r}
set.seed(27)

# only the numeric variables ()
vars <- house[,c(6,7,13:22)]
pc <- prcomp(vars, scale = T)
summary(pc)

sort(round(pc$rotation[,1], 2))

# PCR
pcrMod <- pcr(log_price ~ bedrooms + bathrooms + sqft_living + sqft_lot + floors + 
    waterfront + view + +grade + yr_built + yr_renovated + zipcode + 
    lat + long + sqft_living15 + sqft_above, data = train, ncomp = 22)

rmse(predict(pcrMod, nncomp = 4), train$log_price) # RMSE = 0.3617545
rmse(predict(pcrMod, nncomp = 4), test$log_price) # RMSE = 0.6526302


pcrmse <- RMSEP(pcrMod, newdata = test)
plot(pcrmse, main = "")
which.min(pcrmse$val) # 23 pc
pcrmse$val[23]


pcrMod_2 <- pcr(log_price ~ bedrooms + bathrooms + sqft_living + sqft_lot + floors + 
    waterfront + view + +grade + yr_built + yr_renovated + zipcode + 
    lat + long + sqft_living15 + sqft_above, data = train, validation = "CV", ncomp = 22, segments = 10)
pcrCV <- RMSEP(pcrMod_2, estimate = "CV")
plot(pcrCV, main = "")
```

# Add other models here... 
